
## 1. Installing packages: `feedeR`, `rtweet`, `RCurl`, `xml2`, `RJSONIO`, `RSQLite`, `stringr`, `readxl`

Inside the R prompt, run the following command

```{r}
package_list <- c('feedeR','XML','rtweet','RCurl','xml2','RJSONIO','RSQLite','stringr','readxl')
install.packages(package_list)
```

You can test that all packages are installed and usable by load them all and check session information to see the loaded packages under header *other attached packages*


```{r}
library(feedeR)
library(rtweet)
library(RCurl)
library(XML)
library(xml2)
library(RJSONIO)
library(RSQLite)
library(stringr)
library(readxl)
sessionInfo()
```

## 2. Setup `rtweet` for streaming Twitter data

- This section is based on instructions provided at https://mkearney.github.io/rtweet/articles/auth.html

** 2.1. Creating Twitter account **

- Sign up for a Twitter account at https://twitter.com
- Make sure that your account has an associated phone number. This is required to make your Twitter account into a developer account (being able to create app)

** 2.2. Creating Twitter app **

- Go to https://apps.twitter.com and sign in with your Twitter account
- Create an application:
  - Application names are unique, you will need to pick a different name from R_Workshop_Clemson
  - Some descriptions are required, it is just to describe what your application will do
  - The website is required, but you do not have to provide a specific website. You only need to provide a place holder URL that is in the correct format. 
  - The Callback URL must be http://127.0.0.1:1410
- Once the Twitter app is created, you will be able to click on the app's name on the front page of https://apps.twitter.com to go to the Application Management page. Select the **Keys and Access Tokens** tab to see your access tokens. You will need the **Consumer Key (API Key)** and **Consumer Secret (API Secret)** strings as shown in this tab for the next steps. 

** 2.3. Setting up Twitter security token for R's rtweet package: **

- Continue in the same command line terminal from step 1, type in the following R codes

```{r}
appname <- "R_Workshop_Clemson"
key <- "1EnhV4LwxFJFHms5xBu9H3vvn"
secret <- "7FtYMVQqlO7d1Hd1F8QDkhiuHDYvlhw4gpGhKb4P6Pug3Zicqz"
twitter_token <- create_token(app = appname,
                              consumer_key = key,
                              consumer_secret = secret)
```

- After these R commands are executed, a browser will pop up asking you to sign in and authenticate the access token, and the R environment will print the following lines while waiting for the authentication:

```
Waiting for authentication in browser...
Press Esc/Ctrl + C to abort
```

- Click on the blue **Authorize App** button to confirm the authorization. This will take you to a web page that has the lines *Authentication complete. Please close this page and return to R*. You can now close the Firefox browser. The R environment will print out the line *Authentication complete* and escape from waiting mode into the normal R prompt. 

** 2.4. Saving Twitter security token for future usage: **

The `twitter_token` variable should be saved to a file after step 2.3 is completed so that it can be reused later. The process is as follows:

- Continuing in the same R command line terminal as step 2.3, execute the followings:

```{r}
home_dir <- Sys.getenv('R_USER')
current_dir <- getwd()
cert_dir <- 'cert'

if (!file.exists(cert_dir)){
    dir.create(file.path(current_dir, cert_dir))
} else {
    print ("Directory already exists")
}
  
file_name <- file.path(current_dir, cert_dir, "twitter_token.rds")
saveRDS(twitter_token, file = file_name)
cat(paste0("TWITTER_PAT=", file_name),file = file.path(home_dir, ".Renviron"),append = TRUE)
```

- The above codes will save the `twitter_token` variable into the `twitter_token.rds` file stored inside your home directory. Next, it creates an environment variable called `TWITTER_PAT` that points to this file and store the environment variable in the default `.Renviron` file that will be loaded by R whenever R is started. 
- It is possible to customize the path to the token file (maybe save it to a specific directory) and `TWITTER_PAT`
- When the `rtweet` library is loaded, `TWITTER_PAT` will be read and all subsequent streaming calls to Twitter will be validated automatically. 
