# Learning Objectives

- Understand data acquisition: downloading from static link
- Understand data curation: working with hierarchically structured data (XML)

## Clear environment

```{r}
rm(list = ls())
```

## Where am I?

```{r}
getwd()
```

## Data Curation

- For JSON and XML formats, we will use the [NY Retail Food Store Database from data.gov](https://catalog.data.gov/dataset/retail-food-stores/resource/498a7e81-ea0e-425c-bb8d-a4e36d619f81)


### 2. XML Format

- Extended Markup Language
- Example of data in XML format: `samples/books.xml` from https://msdn.microsoft.com/en-us/library/ms762271(v=vs.85).aspx

Package `XML` reads xml data into a tree structure that can be interpreted by external XML processing functions. 

```{r}
library(xml2)
```

```{r}
sample_xml <- read_xml('./samples/books.xml')
```

```{r}
sample_xml
```

```{r}
str(sample_xml)
```

```{r}
xml_structure(sample_xml)
```

```{r}
book_catalog <- as_list(sample_xml)
```

```{r}
str(book_catalog)
```

```{r}
summary(book_catalog)
```

For XML version of the NY food store data, we don't have separate data and metadata. Instead, everything is stored together as tags, attributes, and values


```{r}
print (getwd())
stores_file <- file.path('data','food_stores.xml') #back up copy in samples/food_stores.xml
stores_url <- 'https://data.ny.gov/api/views/9a8c-vfzj/rows.xml?accessType=DOWNLOAD'
```


```{r}
download.file(stores_url,stores_file, quiet = TRUE)
```


```{r}
stores_xml <- read_xml(stores_file)
```


```{r}
str(stores_xml)
```


```{r}
list_stores <- as_list(stores_xml)
```

Anticipating from the results of `as_list` when called on `book_catalog`, we will not want to risk displaying the structure of a lengthy list. Frequent size checks are recommended


```{r}
print(length(list_stores))
```


```{r}
print(length(list_stores[[1]]))
```


```{r}
print(length(list_stores[[1]][[1]]))
```


```{r}
str(list_stores[[1]][[1]])
```

Similar to the JSON case, to convert elements of this XML-based list into rows of a data frame, we need to first construct the headers:

- XML's tags become names of the list's elements
- XML's attributes become attributes of associated elements within the list


```{r}
names(list_stores[[1]][[1]])
```

We don't want `location` but attributes `latitute` and `longitude` of `location`


```{r}
attributes(list_stores[[1]][[1]][['location']])
```


```{r}
xml_headers <- names(list_stores[[1]][[1]])
location_attributes <- names(attributes(list_stores[[1]][[1]][['location']]))
stores_headers <- xml_headers[1:(length(xml_headers) - 1)]
stores_headers <- c(stores_headers,
                   location_attributes[[2]],
                   location_attributes[[3]])
print(stores_headers)
```

While it is possible to *hardcode* headers information, an implementation that rely on raw data's information is potentially dynamic, maintainable, and reusable


```{r}
stores_counts <- length(list_stores[[1]])

df_stores <- data.frame(character(stores_counts), stringsAsFactors=FALSE)
for (i in 2:14){
    df_stores[,i] <- character(stores_counts)
}

for (i in 15:16){
    df_stores[,i] <- numeric(stores_counts)
}

colnames(df_stores) <- stores_headers
```


```{r}
str(df_stores)
```


```{r}
for (i in 1:1){
    tmpList <- list_stores[[1]][[i]]
    for (j in 1:14){
        if (!is.null(tmpList[[j]]) && length(tmpList[[j]] > 0)){
            df_stores[i,j] <- tmpList[[j]][[1]]
        }
    }
    location_attributes <- attributes(tmpList[['location']])
    for (j in 15:16){
        if (!is.null(location_attributes[[j-13]])){
            df_stores[i,j] <- as.numeric(location_attributes[[j-13]])
        }
    }
    print (df_stores[i,])
}
```


```{r}
for (i in 1:stores_counts){
    tmpList <- list_stores[[1]][[i]]
    for (j in 1:14){
        if (!is.null(tmpList[[j]]) && length(tmpList[[j]] > 0)){
            df_stores[i,j] <- tmpList[[j]][[1]]
        }
    }
    location_attributes <- attributes(tmpList[['location']])
    for (j in 15:16){
        if (!is.null(location_attributes[[j-13]])){
            df_stores[i,j] <- as.numeric(location_attributes[[j-13]])
        }
    }
}
```


```{r}
print(list_stores[[1]][[10]])
print(df_stores[10,])
```

For complex XML data, the recommended approach is to use [XPath Query Language](https://en.wikipedia.org/wiki/XPath):
- /node = top-level node
- //node = node at any level
- node[@attr] = node that has an attribute named "attr"
- node[@attr='something'] = node that has an attribute named "attr" with value 'something'
- node/@attr = value of attribute `attr` in node that has such attributes. 

XPAth queries can be used with package xml2's `xml_path` function to describe operations on specific XML data elements whose tags and attributes match the query patterns